Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-01-18 main.tex
more text

% LTeX: enabled=false
We introduce a specific notion of connectivity when we are given an $n$-dimensional simplicial complex $\calT$. 
We call two $n$-dimensional simplices $S, S' \in \calT$ \emph{face-neighboring} if $S \cap S'$ is a common face of both of them. 
We call $n$-simplices $S,S' \in \calT$ \emph{face-connected in $\calT$} if there exists a sequence $S=S_0,S_1,\dots,S_m=S' \in \calT$ such that $S_{i}$ and $S_{i-1}$ are face-neighboring for all $1 \leq i \leq m$. 
Such a sequence is called a \emph{face path} from $S$ to $S'$ in $\calT$. 
Clearly, face-connected in $\calT$ is an equivalence relation among simplices. 
A \emph{face-connected component} of $\calT$ is an equivalence class under this equivalence relation, 
and we call $\calT$ \emph{face-connected} if it has only one face-connected component. 
% LTeX: enabled=true

@@

        \draw[dotted] (O) -- (A); % THIS ONE 
        \draw[dotted] (O) -- (B); % THIS ONE 
        \draw[] (A) -- (B); % THIS ONE 
%                 \draw (O) -- (C);
%                 \draw (O) -- (D);
%                 \draw (O) -- (E);
        \draw[line width=0.2em] (B) -- (C) -- (D) -- (E) -- (A);
    \end{tikzpicture}%
    \end{tabular} 
    \caption{Sketch of the geometric situation in the proof of Proposition~\ref{proposition:starreflection}.
    The triangle $T$ completes the star around an interior vertex. $A$ is the complement of the triangle within the star. 
    $\Gamma_1$ is the interface between the two sets, and left invariant by the reflection.
    $\Gamma_2$ is left invariant by the deformation.
    $\Xi_{1}$ maps $T$ into $A$ while preserving the interface $\Gamma_1$,
    and 
    $\Xi_{2}$ maps the whole patch into $A$ while preserving the boundary $\Gamma_2$.
    }\label{figure:geometricconstruction}
\end{figure}

@@

        Assuming that $\varrho > 0$ is at most the minimum height of $z_{S}$ in any $n$-simplex $K \in \calK$,
        suppose that $z \in \Ball_{\varrho}(z_S)$. 
        Let $x \in \patch_{\calT}(S)$ be in the interior of the local patch. 
        Suppose that the line segment $I$ from $z$ to $x$ is not wholly contained in $\patch_{\calT}(S)$. 
        Then $I$ intersects with a boundary face $F \subseteq K$ of the local patch.
        The straight line segment from $z_{S}$ to $x$ lies within one open half-space defined by the affine hull of $F$.
        Since $I$ intersects with $F$, this open-half space cannot contain $z$ but it must contain $\Ball_{\varrho}(z_S)$.
        That is a contradiction. Hence, the straight line segment from $z$ to $x$ bust lie within the interior of the local patch. 
        We deduce that $\patch_{\calT}(S)$ is star-shaped with respect to $\Ball_{\varrho}(z_S)$.

@@

We compare the reference Neumann eigenvalues with the gradient potential estimate in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
A brufe-force search easily enumerates all the (reptition-free) face-paths between triangles, which allows us to find the best estimate possible estimate with the aforementioned theorem.

We compare the reference Neumann and Dirichlet eigenvalues with the bounds computed in Theorem~\ref{theorem:poincarefriedrichsestimate:exterior} and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
Here, we take into account Lemma~\ref{lemma:volumecomparison} and Theorem~\ref{proposition:starreflection} (with Remark~\ref{remark:improved_starreflection}).
Again, brute-force search is feasible to enumerate all possible shellings of moderately small two-dimensional triangulations,
allowing us to compute the best possible of these bounds.

@@

The reference Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators are found via standard finite element eigenvalue computations, using the lowest-order finite element de~Rham complex over a sufficiently refined mesh (four steps of uniform refinement).
Again, we compare these reference values with estimates obtained Theorem~\ref{theorem:poincarefriedrichsestimate:exterior} and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}, and in the special case of the gradient, with Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.

The computation becomes more demanding in dimension three. 
Enumerating all the paths between tetrahedra in our example triangulations is feasible but takes considerably longer. 
Brute-force enumeration of all shellings seems infeasible, though. 
Instead,

\todo{How to select shellings}

>>>>

% LTeX: enabled=false
We introduce a specific notion of connectivity when we are given an $n$-dimensional simplicial complex $\calT$. 
We call two $n$-dimensional simplices $S, S' \in \calT$ \emph{face-neighboring} if $S \cap S'$ is a common face of both of them. 
We call $n$-simplices $S,S' \in \calT$ \emph{face-connected in $\calT$} if there exists a sequence $S=S_0,S_1,\dots,S_m=S' \in \calT$ such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$. 
We call such a sequence a \emph{face path} from $S$ to $S'$ in $\calT$. 
Clearly, face-connected in $\calT$ is an equivalence relation among simplices. 
A \emph{face-connected component} of $\calT$ is an equivalence class under this equivalence relation, 
and we call $\calT$ \emph{face-connected} if it has only one face-connected component. 
% LTeX: enabled=true

@@

        \draw[dotted] (O) -- (A); % THIS ONE 
        \draw[dotted] (O) -- (B); % THIS ONE 
        \draw[] (A) -- (B); % THIS ONE 
%                 \draw (O) -- (C);
%                 \draw (O) -- (D);
%                 \draw (O) -- (E);
        \draw[line width=0.2em] (B) -- (C) -- (D) -- (E) -- (A);
    \end{tikzpicture}%
    \end{tabular} 
    \caption{Sketch of the geometric situation in the proof of Proposition~\ref{proposition:starreflection}.
    The triangle $T$ completes the star around an interior vertex. $A$ is the complement of the triangle within the star. 
    $\Gamma_1$ is the interface between the two sets, and left invariant by the reflection.
    $\Gamma_2$ is left invariant by the deformation.}\label{figure:geometricconstruction}
\end{figure}

@@

        Assuming that $\varrho > 0$ is at most the minimum height of $z_{S}$ in any $n$-simplex $K \in \calK$,
        suppose that $z \in \Ball_{\varrho}(z_S)$. 
        Let $x \in \patch_{\calT}(S)$ be in the interior of the local patch. 
        If the line segment $I$ from $z$ to $x$ is not wholly contained in $\patch_{\calT}(S)$, 
        then $I$ passes through a boundary face $F \subseteq K$ of the local patch,
        and $z$ and $x$ are contained on different sides of the hyperplane $H$ spanned by $F$. 
        Then $x$ and $z_{S}$ are the same side of $H$, since otherwise the straight line segment from $x$ to $z_{S}$ would not be contained in the patch. 
        This means, however, that $z_{S}$ and $z$ are on different sides of $H$, contradicting that $\varrho$ is at most the height of $z_S \in K$. 
        % TODO: The above argument seems to rely on convexity. Generalise it to avoid that difficulty.

@@

We compare the reference Neumann eigenvalues with the gradient potential estimate in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
A brufe-force search easily enumerates all the shortest face-paths that start from a given triangle, so we use the best possible estimate that this method provides.

We compare the reference Neumann and Dirichlet eigenvalues with the bounds in Theorem~\ref{theorem:poincarefriedrichsestimate:exterior} and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
Here, we take into account Lemma~\ref{lemma:volumecomparison} and Theorem~\ref{proposition:starreflection} (with Remark~\ref{remark:improved_starreflection}).
Again, brute-force search is feasible to enumerate all possible shellings of moderately small two-dimensional triangulations,
so the best possible estimate with that method can be obtained.

@@

The reference Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators are found via standard finite element eigenvalue computations, using the lowest-order finite element de~Rham complex and a sufficiently refined mesh (four steps of uniform refinement).
As in the two-dimensional case outlined above, we again compare these reference values with estimates obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad}, Theorem~\ref{theorem:poincarefriedrichsestimate:exterior}, and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.

Enumerating all the shortest-path trees of the triangulations above is feasible but takes considerably longer. 
Brute-force enumeration of all shellings seems infeasible, though, due to the combinatorial explosion of the number of possible shellings. 
Instead, we perform a greedy backtracking search: 
starting with a single simplex, we successively increase the number of shellings until a complete shelling is found. 
Among the possible candidate simplices, we prioritize those that introduce the smallest possible geometric constant. 
We moreover prioritize the completion of face stars, since the induced constants are easier to control via Lemma~\ref{lemma:volumecomparison}. 
It is imperative for the improvement of our upper bounds that we compute the relevant geometric parameters, such as heights, individually for each simplex and use individualized estimates for each of the geometric transformations, thus avoiding overtly pessimistic blanket estimates.

This greedy backtracking algorithm does generally not guarantee that the optimal shelling is found before the last iteration. 
Our implementation starts with a fixed simplex and generates up to $10$ shellings. 
We repeat this for a possible choices of initial simplex, and pick the optimal among those.
The candidate shellings that we obtain in this manner allow for estimates that differ by a factor of about $2$.
